**Sequencing 16S rRNA gene fragments using the PacBio SMRT DNA sequencing system**

### 2016-01-30 Schloss_PacBio16S_PeerJ_2016.Rmd
Fix up R code in Conclusion

The widespread adoption of the 454 and MiSeq platforms and decrease in the use of Sanger sequencing for the 16S rRNA gene has resulted in a decrease in the generation of the full-length reference sequences that are needed for performing phylogenetic analyses and designing lineage specific PCR primers and fluorescent *in situ* hybridization (FISH) probes. It remains to be determined whether the error rates we observed for full-length sequences are prohibitive for these applications. We can estimate the distribution of errors assuming that the errors follow a binomial distribution along the length of the 1,500 nt gene with the error rate that we achieved from the V1-V9 mock community data prior to pre-clustering the sequences, which was `r round(100*pooled_error["V1V9","unique"],2)`% (Figure 5). Under these conditions one would only expect `r round(100*dbinom(0, 1500, pooled_error["V1V9","unique"]), 2)`% of the sequences to have no errors and 50% of the sequences would have at least `r qbinom(0.50, 1500, pooled_error["V1V9","unique"])` errors. After applying the pre-clustering denoising step, the erro rate drops by `r round(pooled_error["V1V9","unique"]/pooled_error["V1V9","precluster"])`-fold to `r 100*pooled_error["V1V9","precluster"]`%. With this error rate, we would expect `r round(100*dbinom(0, 1500, pooled_error["V1V9","precluster"]), 2)`% of the sequences to have no sequencing errors. Of course, the tradeoff is the loss of resolution among closely related sequences.

>>>>

```{r binomial_analysis}
unique <- error_summary[error_summary$region=="V1V9" & error_summary$method=="coverage_pred","error"]/100

precluster <- error_summary[error_summary$region=="V1V9" & error_summary$method=="precluster","error"] / 100
unique_to_precluster <- unique / precluster
```

The widespread adoption of the 454 and MiSeq platforms and decrease in the use of Sanger sequencing for the 16S rRNA gene has resulted in a decrease in the generation of the full-length reference sequences that are needed for performing phylogenetic analyses and designing lineage specific PCR primers and fluorescent *in situ* hybridization (FISH) probes. It remains to be determined whether the error rates we observed for full-length sequences are prohibitive for these applications. We can estimate the distribution of errors assuming that the errors follow a binomial distribution along the length of the 1,500 nt gene with the error rate that we achieved from the V1-V9 mock community data prior to pre-clustering the sequences, which was `r round(100*unique,2)`% (Figure 5). Under these conditions one would expect `r round(100*dbinom(0, 1500, unique), 2)`% of the sequences to have no errors and 50% of the sequences would have at least `r qbinom(0.50, 1500, unique)` errors. After applying the pre-clustering denoising step, the error rate drops by `r round(unique_to_precluster, 1)`-fold to `r 100*precluster`%. With this error rate, we would expect `r round(100*dbinom(0, 1500, precluster), 2)`% of the sequences to have no sequencing errors. The cost of the reduced error rate is the loss of resolution among closely related sequences.
